Classical Simulation of Quantum Error Correction in a Fibonacci Anyon Code

### 2015-09-07 fib2.tex
added placeholder for supplemental material.
first pass at addressing ref B's comments

Concretely, we consider the simulation of quantum error correction in a two-dimensional lattice 
system with Fibonacci anyons, a class of nonabelian anyons that are universal for quantum 
computation~\cite{Freedman2002, Nayak2008}. Fibonacci anyons are experimentally motivated as the 
expected excitations of the $\nu=\frac{12}{5}$ fractional quantum Hall 
states~\cite{Slingerland2001}, and can be realized in several spin 
models~\cite{Levin2005, Bonesteel2012, Kapit2013, Palumbo2014} and composite 
heterostructures~\cite{Mong2014}.

We use a flexible phenomenological model of dynamics and thermal 
noise to describe a system with Fibonacci anyon excitations. Within 
this model, we apply existing general topological error-correction protocols, and 
simulate the successful preservation of quantum information encoded in topological 
degrees of freedom. Topological quantum computation protocols using nonabelian anyons 
typically implicitly assume the existence of an error-correction protocol to 
correct for diffusion or unwanted creation of anyons. Our results 
are the first explicit demonstration that such a scheme will 
be successful when applied to a universal topological quantum computer.

@@

@@

In addition to the fusion space there is also a global degeneracy associated with the topology of the 
manifold on which the anyons reside. We consider systems with the topology of a torus, 
which for the Fibonacci anyons gives rise to a 2-fold degeneracy.

@@

Pair-creation events acting within a single tile do not affect the corresponding observable, and so the only pair-creation events we need explicitly consider act between neighboring pairs of tiles. These are then each associated with an edge of the lattice $\Lambda$, chosen 
uniformly at random to model high-temperature, short timescale thermalizing dynamics. The simulation time, and thus the 
error-correction threshold, will be measured in terms of average number of noise processes 
per edge, as opposed to an iid noise probability per edge. The latter measure is not 
appropriate for error processes in nonabelian anyon models, but the two coincide in the limit 
of low error rates in those cases where they are comparable (see Ref.~\cite{Brell2013}).

@@

Below the (bond) percolation threshold for (say) a 2D square lattice, we expect random sets of 
bonds to decompose into separate connected components 
of average size $O(\log(n))$ and variance $O(1)$~\cite{Bazant2000}.
Each noise process in our model is associated with a (randomly distributed) edge, and so 
disconnected components correspond to sets of anyons that could not have interacted at any 
point in their history. 
We are free to neglect the degrees of freedom associated with braiding between components 
because each component has trivial total charge.
This allows us to simulate the braiding processes within each component separately. 
In other words: the quantum state in the fusion space of all anyons factorizes into 
a tensor product over components. 
Since each 
component has size only $O(\log(n))$, we can typically simulate these dynamics efficiently 
because the resulting fusion space has dimension $O(\mathrm{poly}(n))$. 
There are 
random processes that violate this reasoning, but these are suppressed 
exponentially in the lattice size $L$~\cite{Grimmett1989}.

@@

We can guarantee that error-correction will succeed whenever
the action of noise plus decoder 
does not percolate the anyons over the lattice,
but it is possible that percolated events may still result in no error. 
The connection between the percolation threshold and the error-correction threshold 
is not well understood in general~\cite{Hastings2014}, though it is clear 
that our threshold estimate will be a lower bound for 
the true threshold that may be found if all events 
(including those that have percolated) were simulated. 
However, calculations of homologically nontrivial operations for small lattice sizes suggest that almost all such 
processes will indeed result in a logical or leakage error.
As such it is likely that neglecting the full simulation of 
these events does not significantly affect our observed threshold value.

@@

>>>>

Concretely, we consider the simulation of quantum error correction in a two-dimensional lattice 
system with Fibonacci anyons, a class of nonabelian anyons that are universal for quantum 
computation~\cite{Freedman2002, Nayak2008}. Fibonacci anyons are experimentally motivated as the 
expected excitations of the $\nu=\frac{12}{5}$ fractional quantum Hall 
states~\cite{Slingerland2001}, and can be realized in several spin 
models~\cite{Levin2005, Bonesteel2012, Kapit2013, Palumbo2014} and composite 
heterostructures~\cite{Mong2014}.
Any of these kinds of physical systems could be used to perform universal topological quantum computation \cggb{citations?}. Natural sources of noise from thermal fluctuations or external perturbations will be suppressed by the energy gap but must still be corrected to allow for scalable computation.

We use a flexible phenomenological model of dynamics and thermal 
noise to describe a system with Fibonacci anyon excitations. Within 
this model, we apply existing general topological error-correction protocols, and 
simulate the successful preservation of quantum information encoded in topological 
degrees of freedom. Topological quantum computation protocols using nonabelian anyons 
typically implicitly assume the existence of an error-correction protocol to 
correct for diffusion or unwanted creation of anyons. Our results 
are the first explicit demonstration that such a scheme will 
be successful when applied to a universal topological quantum computer.
Although it is not surprising that the standard topological error-correction techniques we use successfully correct errors in universal nonabelian anyon systems, the ability to simulate the precise details of how and when these techniques succeed on finite system sizes has not previously been available. \cggb{Would a comment about the importance of locality in topological codes (compared to abstract QEC) maybe help Ref.~B's confusion about why we don't use ``conventional'' QEC? I'm not totally sure what he wants.}

@@

\cggb{I don't know how to address Ref.~B's comment here.}

@@

In addition to the fusion space there is also a global degeneracy associated with the topology of the 
manifold on which the anyons reside. We consider systems with the topology of a torus, 
which for the Fibonacci anyons gives rise to a 2-fold degeneracy.
This extra degree of freedom can be thought of being associated with the total anyonic charge (or flux) running through the torus itself. Since there are two different possible charges in the Fibonacci anyon model, the global degeneracy is 2.

@@

Pair-creation events acting within a single tile do not affect the corresponding observable, and so the only pair-creation events we need explicitly consider act between neighboring pairs of tiles. These are then each associated with an edge of the lattice $\Lambda$, chosen 
uniformly at random to model high-temperature, short timescale thermalizing dynamics.
In studies of codes based on abelian anyon models, the error correction threshold or memory lifetime is typically quoted in terms of an iid noise strength per edge. This measure is not appropriate in a nonabelian anyon setting, where noise processes on neighbouring edges do not commute. We thus measure the error correction threshold and memory lifetime in terms of the average number of noise processes per edge. 
It is easy to see that, in those cases where they are both appropriate, these two methods coincide in the limit of low error rates (see Ref.~\cite{Brell2013} for more details). The thresholds we consider in this paper are on the order of $0.1$ noise processes per edge. In models where both noise strength measures are appropriate, such as the toric code under bit-flip errors, the discrepancy between the two values is around $10%$ at these strengths.

@@

Below the (bond) percolation threshold for (say) a 2D square lattice, we expect random sets of 
bonds to decompose into separate connected components 
of average size $O(\log(n))$ and variance $O(1)$~\cite{Bazant2000}.
Each noise process in our model is associated with a (randomly distributed) edge, and so 
disconnected components correspond to sets of anyons that could not have interacted at any 
point in their history. 
We are free to neglect the degrees of freedom associated with braiding between components 
because each component has trivial total charge.
This allows us to simulate the braiding processes within each component separately. 
In other words: the quantum state in the fusion space of all anyons factorizes into 
a tensor product over components. 
Since each 
component has size only $O(\log(n))$, we can typically simulate these dynamics efficiently 
because the resulting fusion space has dimension $O(\mathrm{poly}(n))$. 
Of course, this is only a probabilistic statement, and so there may be cases where there exist components with size larger than $O(\log(n))$. However, the likelihood of this is suppressed exponentially in the lattice size $L$~\cite{Grimmett1989}.

@@

We can guarantee that error-correction will succeed whenever
the action of noise plus decoder 
does not percolate the anyons over the lattice,
but it is possible that percolated events may still result in no error. 
The connection between the percolation threshold and the error-correction threshold 
is not well understood in general~\cite{Hastings2014}, though it is clear 
that our threshold estimate will be a lower bound for 
the true threshold that may be found if all events 
(including those that have percolated) were simulated. 
However, calculations of homologically nontrivial operations for small lattice sizes suggest that almost all such 
processes will indeed either cause a logical error or a complete failure to return to the codespace.
As such we expect that neglecting the full simulation of 
these events does not significantly affect our observed threshold value.

@@

\appendix

\section*{Supplementary Material}

\section{Details on the algorithm}
